Reduced growth due to belowground sink limitation is not explained by reduced photosynthesis

### 2016-04-08 manuscript/manuscript_pve.Rmd
moved files out of master (model scenarios).  BM edits

```{r, echo=FALSE}
allomfig <- "1"
airvarsfig <- "2"
allocfig <- "3"
asatfig <- "4"
achemfig <- "5"
massmodelfig <- "6"
massmodelfig2 <- "S1"

@@

Understanding plant growth and its relationship to C assimilation requires knowledge of the mass balance that must be achieved between C uptake and subsequent allocation to growth, storage, and respiration. As woody plants have highly integrated systems of competing carbohydrate sinks [@kozlowski1992carbohydrate], growth should principally depend on the allocation of photosynthate from source leaves among different tissues and organs. At long enough time scales leaf photosynthesis (*A*~n~) and respiratory losses together determine net C balance and will correlate to plant growth. At shorter temporal scales, however, growth can instead be mediated by tissue C storage pools. This discrepancy has led to the current debate on how strongly plant growth is controlled by either source or sink activity. Consequently, plant growth cannot always be simply determined by *A*~n~, making it complex to understand and challenging to model [@fourcaud2008plant]. Despite a wealth of studies, large uncertainties still remain regarding the coordination of C supply and growth of woody species.

@@

##Seedling growth model
We developed a simple seedling growth model that utilized leaf *A*~n~ rates to allocate daily C assimilate towards biomass production of stems, leaves, fine roots and coarse roots. The model begins with mean initial tissue component biomass (leaf~0,i~, stem~0,i~ and root~0,i~) and a starting leaf area (L~0,i~) measured prior to planting. The initial biomass of roots was divided evenly between fine and coarse roots. The daily net biomass production of seedlings (*P*~i~) is then given by

\
(1)
$$P_i(t) = \frac{L_\mathrm{i}~Cday_\mathrm{i}~\sigma - R_\mathrm{i}}{\epsilon}$$

where *L*~i~ is modelled total plant leaf area (m^2^), *Cday*~i~ is the predicted daily carbon assimilation (g C m^-2^ d^-1^, explained below), $\sigma$ is a self-shading parameter (explained below), $\epsilon$ is a biomass conversion efficiency parameter (g C g mass^-1^) and *R*~i~ is the mass based total respiration of all tissue components (g C d^-1^). Here, i represents model runs for each experimental treatment (1:7). Total respiration was calculated as

\
(2)
$$R_i(t) = \Sigma_{c=1}^{3}(R_\mathrm{c}~M_\mathrm{c,i}(t))$$

where *R*~c~ is published or measured tissue respiration rates of different woody tissue components (Table `r pvetabS`, g C g mass^-1^ d^-1^) and *M*~c,i~ is the modelled standing biomass for each component (g). Here, c represents fine roots, coarse roots or stems. Leaf respiration is included in the calculation of *Cday* (described below). The change in individual component biomass (*M*~c,i~), here solved on a daily time step, is given by

\
(3)
$$\frac{dM_\mathrm{c}}{dt} = \theta_\mathrm{c}~P_\mathrm{i}(t)~-~(\Lambda_\mathrm{c}~M_\mathrm{c,i})$$

where $\theta$~c~ is the component specific biomass partitioning to whole plant biomass (%) and $\Lambda$~c~ is component specific turnover rate (yr^-1^). Because we did not observe branch turnover, $\Lambda$ of stems was assumed to equal 0. Total seedling biomass, per time step, was then equal to the sum of all biomass components; leaves, stems, fine roots and coarse roots.

\
*Cday* was predicted by using a coupled photosynthesis - stomatal conductance model [@farquhar1980biochemical;@medlyn2011reconciling] with the 'plantecophys' package [@duursma2015] in R with the mean photosynthetic parameters (*J*~max~, *V*~cmax~, *R* and *g*~1~) for each treatment and meteorological data from an onsite weather station. Values of *J*~max~ and *V*~cmax~ are mean estimates from AC~i~ curves over two measurement periods (explained above), *R* was empirically measured and the *g*~1~ parameter was generated by fitting the optimal stomatal conductance model from [@medlyn2011reconciling] to observed *g*~s~ values. Methods of the coupled leaf gas exchange model are also described in Duursma et al [-@duursma2014peaked]. Combined with the meteorological parameters; PPFD, air temperature, and relative humidity, at 15 min intervals, leaf *A*~n~ rates ($\mu$mol CO~2~ m^-2^ s^-1^) were then predicted for each soil volume treatment. *Cday* was calculated by converting predicted rates to mass C gain over 15 min time steps (g m^-2^) and then summed for 24 h. This resulted in 120 unique values of *Cday* for each soil volume treatment, one value for each day of the experiment. Thus, each daily time step for model runs included a value of *Cday* that represented both treatment specific photosynthetic parameters and meteorological constraints across the duration of the experiment.

\
It was further necessary to calculate a self-shading parameter ($\sigma$) when scaling leaf *A*~n~ with total plant leaf area. This was accomplished by utilizing 61 previously digitized *Eucalyptus* seedlings, covering 5 total species which include *E. tereticornis*, from Duursma et al. [-@duursma2012light] to run in 'YplantQMC' package [@YplantQMC] in R to build a 3D plant structure based on digitized metrics of plant allometry and crown structure. Using the same treatment specific physiological parameters listed above, 'YplantQMC' outputs total *A*~n~, using total leaf area, for seedlings assuming self-shading as well as for a full sun large horizontal leaf. The ratio of total *A*~n~ with self-shading to horizontal leaf was then used to calculate $\sigma$ for each of the 61 digitized seedlings, independently for each of the seven soil volume treatments. Next, the linear relationship between $\sigma$ and total leaf area was determined across digitized seedlings, within each treatment. For the growth model, $\sigma$ was then predicted for each daily time step using the previous days cumulative leaf area and this value  was then applied to *Cday*~i~. All default parameters used in model simulations are reported in Table `r pvetabS`.

\
We then used this model to test the hypothesis that the effects of belowground sink limitation on rates of leaf *A*~n~ were sufficient to accurately predict overall seedling biomass production after 120 days. First, a default model with mean photosynthetic parameters from the free seedlings was optimized to estimate a leaf mass fraction (LMF) which correctly predicted both the leaf mass and total biomass of the harvested free seedling (*M*~free~). This optimized LMF was then used to constrain model runs with treatment-specific *Cday*, while keeping woody tissue respiration and tissue turnover parameters constant, to determine if changes in leaf *A*~n~ alone could predict biomass of seedlings in containers (*M*~pots~). Cumulative net leaf C gain for each treatment was equal to the sum of each value of *Cday*~i~ over 120 days and final seedling C was assumed to equal half of the final dry mass for both modeled and observed seedlings. Seedling biomass production was compared between model output and harvested seedlings with treatment-specific mean *Cday* by first scaling values to the free seedling control.

@@

\
Next, we performed a series of model simulations to test possible C allocation scenarios to account for the over-predictions of seedling C mass. Running the model with observed treatment-specific LMF values from the final harvest (*M*~S1~), which were each lower than the optimized LMF value (see Figure `r allocfig`a), improved model predictions of seedling C mass but still overestimated seedling total C by 32&pm;11.1 % (Figure `r massmodelfig2`a). Using harvest values of LMF, however, does not capture the observed increase in leaf turnover of seedlings in small containers (Figure `r allomfig`c). Thus, the use of harvest LMF values for seedlings in containers in *M*~S1~ likely underestimates daily C allocation to leaves over the final months of the experiment. Increases of 50 % in non-leaf tissue respiration (*M*~S2~) improved biomass estimates slightly but nonetheless overestimated mass C by an average of 46&pm;9.3 % in seedlings with soil volume restriction (Figure `r massmodelfig2`b). With *M*~S2~, non-tissue respiration rates would need to be increased by ca. 250% in order for mass balance to be achieved.

@@

\
Increasing rates of non-leaf respiration (*M*~S2~) improved biomass predictions but to a far lesser extent than changes to leaf C allocation. We also show that very large increases in non-leaf respiration rates would have been required to accurately predict observed seedling biomass. Although the fraction of photosynthate used in respiration is known to vary among species and is sensitive to changes in growth rates [@Lambers2008], results from *M*~S2~ highlight a lack of knowledge regarding how respiration rates of individual tissues, within a single plant, maybe be differentially affected by environmental change. This is noteworthy, as C balance is a delicate equilibrium between fluxes of *A*~n~ and respiration, partial accounting of C dynamics can easily lead to erroneous conclusions [@valentini2000respiration]. These results imply that using fixed rates of respiration in models likely underestimates plant responses to sink limitation. Thus, we agree with Delucia et al. [-@delucia2007forest] that it is likely inappropriate to assume that respiration is a constant fraction of gross primary production in models. Our findings reveal that a combination of different mechanisms, beyond *A*~n~, are likely at play in driving the observed seedling biomass response to sink manipulation. However, the degree to which these mechanisms will regulate growth will undoubtedly shift across different experimental manipulations and plant species.

@@

```{r model_scenarios, fig.width=6, fig.height=8, echo=FALSE, dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/massmodel_optimize_scenarios.R")
```
**Figure S1**. Sensitivity testing of the seedling growth model to different carbon allocation strategies, including; (a) leaf mass fraction set to treatment-specific final harvest values and (b) increases in respiration of non-leaf tissue components by 50 %.  Open and filled symbols represent default model and harvest values, respectively, while shaded symbols represent model sensitivity to each scenario by soil volume treatment. Both seedling carbon mass and daily carbon assimilation were first scaled to the free seedling control.

>>>>

```{r, echo=FALSE}
allomfig <- "1"
airvarsfig <- "2"
allocfig <- "3"
asatfig <- "4"
achemfig <- "5"
massmodelfig <- "6"

@@

Understanding plant growth and its relationship to C assimilation requires knowledge of the mass balance that must be achieved between C uptake and subsequent allocation to growth, storage, and respiration. As woody plants have highly integrated systems of competing carbohydrate sinks [@kozlowski1992carbohydrate], growth is driven by the allocation of photosynthate from source leaves among these different tissues sinks. At long enough time scales leaf photosynthesis (*A*~n~) and respiratory losses together determine net C balance and plant growth. At shorter temporal scales growth can instead be mediated by tissue C storage pools [@chapin1990ecology], disrupting the coordination between *A*~n~ and growth. This discrepancy has led to the current debate on how strongly plant growth is controlled by either source or sink activity.

As plant growth cannot simply be determined by *A*~n~, growth models must take into account several simultanesous processes affecting the C balance at any given time [@fourcaud2008plant].

Despite a wealth of studies, large uncertainties still remain regarding the coordination of C supply and growth of woody species.
*when growth is limited* interpretations of how the C balance reponds.......

@@

##Seedling growth model
We developed a simple seedling growth model that utilized leaf *A*~n~ rates to allocate daily C assimilate towards biomass production of stems, leaves, fine roots and coarse roots. The model begins with mean initial tissue component biomass (leaf~0, stem~0~ and root~0~) and a starting leaf area (L~0~) measured prior to planting. The initial biomass of roots was divided evenly between fine and coarse roots. The daily net biomass production of seedlings (*G*), from @makela1997carbon, is then given by

\
(1)
$$G(t) = \frac{L~P~\sigma - R}{\epsilon}$$

where *L* is modelled total plant leaf area (m^2^), *P* is the daily carbon assimilation per unit leaf area (g C m^-2^ d^-1^, explained below), $\sigma$ is a self-shading parameter (explained below), $\epsilon$ is a biomass conversion efficiency parameter (g C g mass^-1^) and *R* is the mass based total respiration of all tissue components (g C d^-1^). Total respiration was calculated as

\
(2)
$$R(t) = R_\mathrm{s}M_\mathrm{s}~+~R_\mathrm{fr}M_\mathrm{fr}~+~R_\mathrm{cr}M_\mathrm{cr}$$

where *R* is published or measured tissue respiration rates (g C g mass^-1^ d^-1^) of stems (s), fine roots (fr) and coarse roots (cr) and *M* is the modelled standing biomass for each tissue component (g). Leaf respiration is included in the calculation of *P*. The change in each tissue component biomass (*M*) on daily time steps, including leaves, stems fine roots and coarse roots, is given by

\
(3)
$$\frac{dM}{dt} = \theta~P(t)~-~(\Lambda~M)$$

where $\theta$ is the component specific biomass partitioning to whole plant biomass (%) and $\Lambda$ is component specific turnover rate (yr^-1^). Because we did not observe branch turnover, $\Lambda$ of stems was assumed to equal 0. Total seedling biomass, per time step, was then equal to the sum of all biomass components; leaves, stems, fine roots and coarse roots.

##Model parameters and simulations
*P* was predicted by using a coupled photosynthesis - stomatal conductance model [@farquhar1980biochemical;@medlyn2011reconciling] with the 'plantecophys' package [@duursma2015] in R with the mean photosynthetic parameters (*J*~max~, *V*~cmax~, *R* and *g*~1~) for each treatment and meteorological data from an onsite weather station. Values of *J*~max~ and *V*~cmax~ are mean estimates from AC~i~ curves over two measurement periods (explained above), *R* was empirically measured and the *g*~1~ parameter was generated by fitting the optimal stomatal conductance model from [@medlyn2011reconciling] to observed *g*~s~ values. Methods of the coupled leaf gas exchange model are also described in Duursma et al [-@duursma2014peaked]. Combined with the meteorological parameters; PPFD, air temperature, and relative humidity, at 15 min intervals, leaf *A*~n~ rates ($\mu$mol CO~2~ m^-2^ s^-1^) were then predicted for each soil volume treatment. *P* was calculated by converting predicted rates to mass C gain over 15 min time steps (g m^-2^) and then summed for 24 h. This resulted in 120 unique values of *P* for each soil volume treatment, one value for each day of the experiment. Thus, each daily time step for model runs included a value of *P* that represented both treatment specific photosynthetic parameters and meteorological constraints across the duration of the experiment.

\
It was further necessary to calculate a self-shading parameter ($\sigma$) when scaling leaf *A*~n~ with total plant leaf area. This was accomplished by utilizing 61 previously digitized *Eucalyptus* seedlings, covering 5 total species which include *E. tereticornis*, from Duursma et al. [-@duursma2012light] to run in 'YplantQMC' package [@YplantQMC] in R to build a 3D plant structure based on digitized metrics of plant allometry and crown structure. Using the same treatment specific physiological parameters listed above, 'YplantQMC' outputs total *A*~n~, using total leaf area, for seedlings assuming self-shading as well as for a full sun large horizontal leaf. The ratio of total *A*~n~ with self-shading to horizontal leaf was then used to calculate $\sigma$ for each of the 61 digitized seedlings, independently for each of the seven soil volume treatments. Next, the linear relationship between $\sigma$ and total leaf area was determined across digitized seedlings, within each treatment. For the growth model, $\sigma$ was then predicted for each daily time step using the previous days cumulative leaf area and this value was then applied to *P*. All default parameters used in model simulations are reported in Table `r pvetabS`.

\
We then used this model to test the hypothesis that the effects of belowground sink limitation on rates of leaf *A*~n~ were sufficient to accurately predict overall seedling biomass production after 120 days. First, a default model with mean photosynthetic parameters from the free seedlings was optimized to estimate a leaf mass fraction (LMF) which correctly predicted both the leaf mass and total biomass of the harvested free seedling (*M*~free~). This optimized LMF was then used to constrain model runs with treatment-specific *P*, while keeping woody tissue respiration and tissue turnover parameters constant, to determine if changes in leaf *A*~n~ alone could predict biomass of seedlings in containers (*M*~pots~). Cumulative net leaf C gain for each treatment was equal to the sum of each value of *P* over 120 days and final seedling C was assumed to equal half of the final dry mass for both modeled and observed seedlings. Seedling biomass production was compared between model output and harvested seedlings with treatment-specific mean *P* by first scaling values to the free seedling control.

@@

@@

\
**this section needs to address why different, implications for models (what they do and need to do)*

Although the fraction of photosynthate used in respiration is known to vary among species and is sensitive to changes in growth rates [@Lambers2008], it is often conisdered a static parameter in process based growth models.

These results imply that using fixed rates of respiration in models likely underestimates plant responses to sink limitation. Thus, we agree with Delucia et al. [-@delucia2007forest] that it is likely inappropriate to assume that respiration is a constant fraction of gross primary production in models.

A lack of knowledge regarding how respiration rates of individual tissues, within a single plant, maybe be differentially affected by environmental change.

This is noteworthy, as C balance is a delicate equilibrium between fluxes of *A*~n~ and respiration, partial accounting of C dynamics can easily lead to erroneous conclusions [@valentini2000respiration].

Our findings reveal that a combination of different mechanisms, beyond *A*~n~, are likely at play in driving the observed seedling biomass response to sink manipulation. However, the degree to which these mechanisms will regulate growth will undoubtedly shift across different experimental manipulations and plant species.

Empirical data in this study point to *A*~n~ driving reductions in growth, however, modelling result highlight the importance of investigating other components of C balance.

*growth limitation affects both photosynthesis and other components of the C balance*

@@
